\section{Summary} \label{sec:summary}

We have presented a parallel algorithm for interpolation and introduced a novel parallel
algorithm for spreading forces from time-dependent interfaces in the immersed boundary
method. These algorithms have a runtime that is independent of the Eulerian grid. This
makes the parallelization useful for cases where IB points inhabit only a small
percentage of the Eulerian grid cells. We have also introduced two variants for spread
which trade off dependence on the Eulerian grid, in the form of a few vector adds and
memory allocation, for improved runtimes for small enough grids.

These algorithms exhibit nearly ideal scaling, on both the CPU and GPU, for problems of
fixed size and increasing number of threads, as well as for problems of increasing size
and number of threads.  We observed that on the GPU, larger problem sizes led to higher
speedup plateaus, indicating that larger problems on more capable hardware will achieve
even larger speedups than presented here.

Our work does not consider how the algorithms will perform on multi-node systems, though
we expect them to perform well, as the algorithms do not depend on partitioning the
Eulerian or Lagrangian points. Because information about the Eulerian grid is encoded by
the functions $\key$ and $\#$, a distributed-memory algorithm may use a different $\key$
or $\#$ for each node. For other considerations on load-balancing for IB calculations on
distributed-memory systems, we refer the reader to Griffith, \textit{et al.}
\cite{Griffith:2007uk}. As simulations become more computationally intensive, such
algorithms will become more necessary. To that end, we will adapt these algorithms for
multi-node and multi-GPU systems.

% vim: cc=90 tw=89
