\section{Review of the immersed boundary method}

Consider a $d$-dimensional ($d=2$ or 3) rectangular domain $\domain$, which is filled
with a viscous incompressible fluid with constant viscosity $\mu$ and density $\rho$, and
contains an immersed elastic structure, $\interface$. The structure is impermeable to the
fluid and moves at the local fluid velocity, is deformed by this motion, and imparts a
force to the fluid. Otherwise, the interface is treated as part of the fluid. 

The fluid velocity, $\u = \u(\x,\,t)$, and pressure, $p = p(\x,\,t)$, are governed by the
incompressible Navier-Stokes equations for a Newtonian fluid,
\begin{gather}
    \label{eq:ins-evol}
    \rho(\u_t + \u\cdot\grad\u) = \mu\Delta\u - \grad p + \f, \\
    \label{eq:ins-incomp}
    \div\u = 0,
\end{gather}
where $p$ is the pressure and $\f$ is the elastic force density. This is a set of $d+1$
equations in $d+1$ unknowns: the $d$ components of $\u$, and $p$. The equations are
written relative to the Eulerian frame, so that the coordinates $\x$ are independent
variables. Throughout, we write Eulerian quantities in the lowercase Latin alphabet.

Let $\X=\X(\params,\,t)$ represent a parametrization of the Cartesian coordinates of the
immersed interface with material coordinates $\params$ at time $t$. Let $\L[\X]$ be the
energy density functional for the elastic interface material. The elastic force density
is computed by evaluating the Fréchet derivative of $\L$,
\begin{equation}
    \label{eq:forces}
    \F = -\delta \L[\X],
\end{equation}
where $\delta$ represents the first variation. Uppercase Latin letters represent
Lagrangian quantities and are functions of $\params$ and $t$.

The interface moves at the local fluid velocity, and force balance on the interface
between the interface and fluid dictates that the interface force on the fluid is equal
to the elastic force. Analytically, the fluid-interface interactions can be written
\begin{gather}
    \label{eq:interpolation}
    \U(\params,\,t) = \int_\domain \Dirac(\x-\X(\params,\,t)) \u(\x,\,t)\d\x,\ \text{and} \\
    \label{eq:spreading}
    \f(\x,\,t) = \int_\interface \Dirac(\x-\X(\params,\,t))\F(\params,\,t)\d\params,
\end{gather}
where $\U(\params,\,t)$ represents the derivative of $\X(\params,\,t)$ with respect to
$t$, and $\Dirac(\x-\X(\params,\,t))$ is the Dirac $\Dirac$-function centered at
$\X(\params,\,t)$.  Equation \eqref{eq:interpolation} is called \term{interpolation}, and
the result of the right-hand side is the fluid velocity at $\X(\params,\,t)$; namely,
$\U(\params,\,t) = \u(\X(\params,\,t),\,t)$.  Equation \eqref{eq:spreading} is called
\term{spreading}, because while $\F(\params,\,t)$ has units of force per unit \emph{area}
on $\interface$, $\f(\x,\,t)$ has units of force per unit \emph{volume} in $\domain$. The
force $\F(\params,\,t)\d\params$ over area $\d\params$ is ``spread'' to the force
$\f(\x,\,t)\d\x$ over volume $\d\x$. 

The fluid equations \eqref{eq:ins-evol}--\eqref{eq:ins-incomp} are spatially discretized
on a regular grid of spacing $h$ so that $\domain$ is divided into square or cubic cells
of side length $h$. Because of the checkerboard instability (see, e.g.,
\cite{Wesseling:2001ci}) in solving the Navier-Stokes equations on collocated regular
grids, we stagger the components of Eulerian vector quantities. As a result, different
components are discretized at different locations, but corresponding components are
discretized on the same grid. Grid cells for these staggered grids do not coincide. We
therefore consider each of these grids individually. We use $\domain_h$ to denote the set
of grid points for the grid under consideration, and $n_\omega$ to denote the number of
grid points. The momentum equation \eqref{eq:ins-evol} is discretized in time using a
second-order implicit-explicit Runge-Kutta method \cite{Peskin:2002go}, and the
incompressibility condition \eqref{eq:ins-incomp} is satisfied using PmIII
\cite{Brown:2001bq}.

The Lagrangian force density \ref{eq:forces} is evaluated at a set of points, usually a
\emph{fixed} set of points in material coordinates, referred to as IB points. The
notation $\X_j=\X(\params_j,\,t)$ refers to an individual IB point on $\interface$ in
Cartesian coordinates. The typical heuristic for distributing the points $\X_j$ on a
connected interface places neighboring IB points at most $h$ apart from one another, and
often at most $h/2$ apart. We therefore denote the set of IB points by $\interface_h$ and
the number of IB points by $n_\gamma$. From these points, we construct a smooth
approximation to the interface using radial basis functions \cite{Shankar:2015km}. This
allows us to calculate geometric properties using well-known formulas, and evaluate the
forces \eqref{eq:forces} analytically \cite{Maxian:2018ek}.

The singular integrals \eqref{eq:interpolation} and \eqref{eq:spreading} do not lend
themselves easily to evaluation. In particular, it is unlikely that IB points and grid
points will coincide. For a regular grid with spacing $h$, we replace the Dirac
$\Dirac$-function with a regularized kernel, $\Dirac_h$, which is a product of
one-dimensional kernels, $h^{-1}\kernel(h^{-1}x)$. There are several options for
$\kernel$ \cite{Griffith:2020hi}, but we will restrict ourselves to the simple cosine
kernel \cite{Peskin:2002go}.

Let $\domain_h$ be one of the Eulerian grids for vector-valued quantities. Discretizing
equations \eqref{eq:scalar-interp} and \eqref{eq:scalar-spread} on $\domain_h$ and
$\interface_h$, respectively, yields
\begin{align}
    \label{eq:disc-interp}
    \dot{X}^n_j &= \sum_i \delta_h(\x_i-\X^n_j)u^n_i h^d \quad \text{and} \\
    \label{eq:disc-spread}
    f^n_i &= \sum_j \delta_h(\x_i-\X^n_j)F^n_j \d\theta_j,
\end{align}
where $u^n_i$ and $f^n_i$ are discrete approximations to the components of $\u$ and $\f$
at $\x_i\in\domain_h$ and time $t=t_n$, and $\dot{X}^n_{\smash j}$ and $F^n_{\smash j}$
are their Lagrangian counterparts at $\X_{\smash j}$, respectively. The terms $h^d$ and
$\d\theta_{\smash j}$ are integration weights analogous to $\d\x$ and $\d\params$. Each
of the above equations look like a matrix-vector multiplication, so we define the
\emph{spreading matrix} $\spread=(\delta_h(\x_i-\X^n_{\smash j}))$ with row $i$ and
column $j$. We call its transpose, $\interp$, the \emph{interpolation matrix}. Collecting
values $u^n_ih^d$ as $\u^n$ at each grid point and $F^n_{\smash j}\d\theta_{\smash j}$ as
$\F^n$ at each IB point, we rewrite Equations \eqref{eq:disc-interp} and
\eqref{eq:disc-spread} in matrix form as
\begin{align}
    \label{eq:matrix-interp}
    \dot{\X}^n &= \interp\u^n \quad\text{and} \\
    \label{eq:matrix-spread}
    \f^n &= \spread\F^n,
\end{align}
respectively.

Collectively, the equations \eqref{eq:ins-evol}--\eqref{eq:spreading} constitute the
IB method, and while different implementations depend on myriad details, a single step
proceeds with timestep $k$ roughly as follows:
\begin{enumerate}[label=(\texttt{\alph*})]
    \item interpolate $\u^n$ to $\X^n$ to get $\U^\ast$,
    \item predict Lagrangian positions $\X^\ast = \X^n + k\U^\ast$,
    \item compute Lagrangian forces $\F^\ast$ using positions $\X^\ast$,
    \item spread $-\F^\ast$ from $\X^\ast$ to get $\f^\ast$,
    \item solve for updated Eulerian velocities $\u^\ast$,
    \item project $\u^\ast$ into space of divergence-free vector fields to get
        $\u^{n+1}$,
    \item interpolate $\u^{n+1}$ to $\X^n$ to get $\U^{n+1}$, and
    \item update $\X^{n+1} = \X^n + k \U^{n+1}$.
\end{enumerate}
We group these steps into 3 categories: the purely Eulerian (\texttt{e}) and
(\texttt{f}); the purely Lagrangian (\texttt{b}), (\texttt{c}), and (\texttt{h}); and the
Eulerian-Lagrangian coupling (\texttt{a}), (\texttt{d}), and (\texttt{g}). We discuss the
%parallelization of the
first two categories \textcolor{red}{in a forthcoming paper}.
%
%briefly before tackling the third.
%
%\section{Parallelization in the Eulerian and Lagrangian frames}
%
%Given an Eulerian force density $\f$ and Lagrangian fluid velocity $\U$ it is
%a straightforward matter to parallelize the evolution of the fluid velocity and interface
%forces, as they are otherwise independent of one another. Assuming fixed sets of grid
%points and material coordinates for IB points, any interdependence of points within a
%frame can be accounted for at the outset to ease parallelization during the course of a
%simulation.
%
%To solve for the fluid velocity, we use the 2nd-order Runge-Kutta method described in
%\cite{Peskin:2002go}.  We compute the advection term in conservation form. For, e.g., the
%quantity $(uv)_y$, we average $u$ in the $y$ direction and $v$ in the $x$ direction. The
%product of these averages is then differentiated in the $y$ direction. We correct near
%boundaries to ensure exact recovery of parabolic flows. This involves scaling certain
%terms by a constant factor to account for errors in discretizing the Laplacian near the
%boundary. To take advantage of the GPU, the solver must be highly parallel. To that end,
%the Helmholtz equations that arise from discretizing equation \eqref{eq:ins-evol} are
%solved using conjugate gradients preconditioned with Chebyshev iteration. Chebyshev
%iteration is an attractive method because, other than parallel vector addition and
%scaling, it requires only the ability to perform sparse matrix-vector multiplication.
%This can be achieved with the appropriate sparse storage formate and a linear algebra
%library such as Intel's \texttt{MKL} or NVIDIA's \texttt{cuSPARSE}. Incompressibility is
%enforced using PmIII of \cite{Brown:2001bq}.  Satisfying the incompressibility condition
%\eqref{eq:ins-incomp} requires a Poisson solve, which is done using conjugate gradients
%preconditioned with full-weighting multigrid. The multigrid routine also uses Chebyshev
%iteration, restricted to high-frequency modes, for relaxation. The resulting solver is
%highly parallel and is suitable for any combination of boundary conditions.
%
%Elastic structures are modeled using radial basis functions (RBFs) to construct the
%surface, as in \cite{Shankar:2015km}. The Fr\'echet derivative in equation
%\eqref{eq:forces} is computed analytically, following \cite{Maxian:2018ek}, and evaluated
%by computing tangents and higher-order surface derivatives on the structure. These
%differential operators are dense, but computed only once and used throughout the course
%of a simulation. The forces are used in equation \eqref{eq:spreading}, for which we
%require integration weights. As the structures of interest are topologically spheres, we
%follow \cite{Fuselier:2013coba} to compute weights on the sphere, and geometric
%information about both the sphere and the structure to obtain appropriate weights for the
%structure \cite{Maxian:2018ek}. The construction of the operators and their application
%requires a way to solve linear systems and perform dense matrix-matrix and matrix-vector
%multiplications. The \texttt{LAPACK} and \texttt{BLAS} libraries are the standard. There
%are numerous parallel implementations of these libraries; for the GPU, we employ NVIDIA's
%implementations, \texttt{cuSOLVER} and \texttt{cuBLAS}.
%
%The interaction operations, however, must take into account the motion of the Eulerian
%and Lagrangian frames relative to one another, and must therefore be recomputed every
%timestep. The remainder of this paper discusses the implementation and parallelization
%of these operations.
The remainder of this paper discusses the implementation and parallelization of the
third.

% vim: cc=90 tw=89
